Implementation and assessment of tools for integrating forest roads into projections of the cumulative effects of disturbance on wildlife

### 2024-05-14 analysis/paper/paper.Rmd
added ilcp gdistance to supp, trying to fix fig 3 size WIP

Road network data was obtained by combining the digital road atlas [@data_bc_digital_2021] and forest tenure road [@british_columbia_ministry_of_forests_lands_natural_resource_operations_and_rural_development_forest_2021] data sets for British Columbia. To avoid situations where the same road was included in both data sets with slightly different coordinates, we removed all roads from the digital road atlas that were within 50 m of forest tenure roads. The existing road network, used as the starting point for projections, was created by filtering out forest tenure roads with an award date greater than 1990. We assumed everything in the digital road atlas was built before 1990. Cutblock information was obtained from the Harvest Areas of BC (Consolidated Cutblocks) dataset [@government_of_british_columbia_harvested_2021], compiled from provincial Forest Cover, the RESULTS Reporting system, Forest Tenure applications, and satellite imagery using change detection processes. Forest harvest cutblocks with harvest years greater than 1990 were filtered and used as targets for the projections. We further filtered cutblocks that contain observed roads from those that do not to reduce errors associated with open areas that have been misidentified as cutblocks.

@@

```{r proj-meth-fig, fig.height=8, fig.width=fig_widths["two"], fig.cap="Comparison of road projection method variants (Table 2). The large four sided cutblock in the top right was added to more clearly show algorithm behaviour within cutblocks.The grade penalty method for setting edge weights (top row) encourages roads to follow slope contours, and yields fewer obviously unrealistic 90 degree turns in steep terrain than our simple cost method (middle row). Selecting a single landing at the centroid of each cutblock (bottom left) creates too few roads within cutblocks. In this 1 ha resolution example the Hardy QGIS plugin (bottom middle) creates more curved roads than our simple cost method (middle row), and does not follow slope contours like the grade penalty method (top row). Landing density is high for all except the centroid and Hardy variants."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.tiff"))
```

@@

Examination of spatial mismatch between observed and projected roads also highlighted errors in our input data. We note these errors because they are not unique to our study area; there is considerable variation in the consistency and accuracy of available resource road information across Canada and elsewhere [@poley_identifying_2022; @british_columbia_ministry_of_forests_lands_natural_resource_operations_and_rural_development_resource_nodate] that users should keep in mind interpreting discrepancies between observed and projected road networks.

>>>>

Road network data was obtained by combining the digital road atlas [@data_bc_digital_2021] and forest tenure road [@british_columbia_ministry_of_forests_lands_natural_resource_operations_and_rural_development__flnrord_forest_2021] data sets for British Columbia. To avoid situations where the same road was included in both data sets with slightly different coordinates, we removed all roads from the digital road atlas that were within 50 m of forest tenure roads. The existing road network, used as the starting point for projections, was created by filtering out forest tenure roads with an award date greater than 1990. We assumed everything in the digital road atlas was built before 1990. Cutblock information was obtained from the Harvest Areas of BC (Consolidated Cutblocks) dataset [@government_of_british_columbia_harvested_2021], compiled from provincial Forest Cover, the RESULTS Reporting system, Forest Tenure applications, and satellite imagery using change detection processes. Forest harvest cutblocks with harvest years greater than 1990 were filtered and used as targets for the projections. We further filtered cutblocks that contain observed roads from those that do not to reduce errors associated with open areas that have been misidentified as cutblocks.

@@

```{r proj-meth-fig, fig.height=9.33, fig.width=fig_widths["two"], out.width="100%", out.height="100%", fig.cap="Comparison of road projection method variants (Table 2). The large four sided cutblock in the top right was added to more clearly show algorithm behaviour within cutblocks.The grade penalty method for setting edge weights (top row) encourages roads to follow slope contours, and yields fewer obviously unrealistic 90 degree turns in steep terrain than our simple cost method (middle row). Selecting a single landing at the centroid of each cutblock (bottom left) creates too few roads within cutblocks. In this 1 ha resolution example the Hardy QGIS plugin (bottom middle) creates more curved roads than our simple cost method (middle row), and does not follow slope contours like the grade penalty method (top row). Landing density is high for all except the centroid and Hardy variants."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.tiff"), dpi=NA)
```

@@

Examination of spatial mismatch between observed and projected roads also highlighted errors in our input data. We note these errors because they are not unique to our study area; there is considerable variation in the consistency and accuracy of available resource road information across Canada and elsewhere [@poley_identifying_2022; @ministry_of_forests_resource_nodate] that users should keep in mind interpreting discrepancies between observed and projected road networks.

### 2024-05-14 analysis/paper/supplement.Rmd
added ilcp gdistance to supp, trying to fix fig 3 size WIP

>>>>

```{r profiling-ilcp_gdistance, fig.cap="Profiling results for the ILCP method using the gdistance package internally showing functions that take > 1 s to run on a cost raster with 1 million cells. Functions higher in the call stack are called by functions lower down.", fig.height=5}
profstat(here::here("analysis/data/derived_data/ilcp_gdistance_100_Rprof.out"))
```
